---
title: "source.rmd"
author: "Nirvikalpa Projects"
date: "8 June 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Source

This application depends on [Open Weather Map](https://home.openweathermap.org/) for its current weather across Indian Terrain.

Open Weather Map provides the API key and with the help of "owmr" package, one can directly download the data from Open Weather from the R console.

Check out the [owmr package](https://cran.r-project.org/web/packages/owmr/owmr.pdf). 

The application also depends on the data for major indian cities over last few years and some cases over the past one century.

Rainfall Data has been gathered from [kaggle](https://www.kaggle.com/rajanand/rainfall-in-india/data). Kaggle is amazing platform for data science enthusiasts and professional. 

One of the major source of data has been National Oceanic & Atmospheric Adminstration. Do check their website [here](http://www.noaa.gov/)

Data can for any category can be obtained from their [climate data portal](https://www7.ncdc.noaa.gov/CDO/country). 

Now I will share the source code for this application which i have developed in last 48 hours. You may use the below source code and make more complex and efficient application out of this.

The souce code is also available on [github](https://github.com/richie312/Weather_App_Indian_Sub-terain)

Below is the source code for shiny application

## Source Code

#### UI.R

```{r,eval=FALSE}
library(shiny)
library(shinythemes)
library(leaflet)
library(owmr)
library(ggplot2)
library(gganimate)
library(plyr)
library(dplyr)
library(yarrr)
library(ggmap)
shinyUI(fluidPage(theme = shinytheme('cerulean'),
                  
                  navbarPage(  title = "Indian Terrain Climate",  
                               id= "nav",
                               tabPanel("Climate", value ="Climate",
                                        ## First Row
                                        
                                        fluidRow(style="background-color:white;",
                                                 column(6,column(12,style="background-color:#239eba;",
                                                                 helpText(h2("Current Weather",style="color:white;")),
                                                                 div(style = "height:50px;")),
                                                        helpText(h5("Data Source:National Oceanic and Atmospheric Administration(US)
                                                                    & https://www.openweathermap.org")),
                                                        
                                                        leafletOutput("map"),
                                                       div(style = "height:20px;"))
                                                          , 
                                                          column(6, column(12,style="background-color:#239eba;",
                                                                           helpText(h4("Historical Data",style="color:white;")),
                                                                           
                                                                 column(5,selectInput(inputId="station", label = "Weather Station", 
                                                                                      choices = c("Leh",
                                                                                                  "Bangalore",
                                                                                                  "Gangtok",
                                                                                                  "Dehradun",
                                                                                                  "Kolkata",
                                                                                                  "Bikaner",
                                                                                                  "Srinagar"),
                                                                                      selected='Gangtok')),
                                                                 column(4,textInput(inputId="color",value = "#b6e5af",
                                                                                    label="Color Code"))),
                                                                 imageOutput("plot1"),div(style = "height:20px;"))),
                                        
                                        ## Next Row
                                        
                                                 fluidRow(style="background-color:white;",
                                                          column(6,
                                                                 column(12,style="background-color:#239eba;",
                                                                 helpText(h4("Historical Rainfall Data for past 115 years",style="color:white;")),
                                                                 column(3,radioButtons(inputId="Rainfall",label="Rainfall",
                                                                              choices = c("pre 1950","post 1950"),
                                                                              selected="post 1950")),
                                                                 column(3,selectInput(inputId="Season",label="Season",
                                                                             choices=c("Winter","Summer","Monsoon","Fall")))),
                                                        plotOutput("yarrplot1"), div(style = "height:600px;")
                                                                 
                                                 ),
                                                 column(6,
                                                        
                                                        column(12,style="background-color:#239eba;",
                                                               helpText(h4("Surface Temperature for past 4 years",style="color:white;")),
                                                               column(6, selectInput(inputId="Temperature",label="Average Temperature",
                                                                                     choices=c("Above Mean Temp of 4 years", "Below Mean Temp of 4 years"))),
                                                               div(style = "height:50px;")),
                                                               
                                                        helpText(h4("Data Source: National Oceanic & Atmospheric Administration",style="color:#239eba;")),
                                                        
                                                        imageOutput("plot2")
                                                        
                                                        )
                                        )
                               ),
                               ## Next Panel
                               
                               tabPanel("Data and Source Code",value = "Resources",
                                        fluidRow(column(12,style="background-color:#D5F5E3;"
                                                        ))
                                        
                                        
                               )
                     
                  )
 )
)



```

#### server.R

```{r,eval = FALSE}
library(shiny)
library(shinythemes)
library(leaflet)
library(owmr)
library(ggplot2)
library(gganimate)
library(plyr)
library(ggthemes)
library(dplyr)
library(yarrr)
library(ggmap)


## Set the API key with Open Weather Map

API_Key<-owmr_settings("Your Own API Key")


server<-shinyServer(function(input,output){
  
  output$map<-renderLeaflet({
    
    
    ## Get the longitude and latitude for Indian cities
    
    India<-owm_cities[owm_cities$countryCode=="IN",]
    India_unique<-India[!duplicated(India[,2]),]
    India_unique<-India_unique[,c(2,3,4)]
    
    
    ## List of my favourite cities
    
    favourite_cities<-c("Leh","Bikaner","Jaisalmer","Ajmer",
                        "Manali","Darjiling","Gangtok","Varanasi","Kolkata",
                        "Guwahati","Bangalore","Chennai","Mumbai","Pune",
                        "Mirik","Rishikesh","Kargil")
    get_city<-function(i){
      x<-India_unique[India_unique$nm == favourite_cities[i],]
      return(x)
    }
    
    India_Fav<-lapply(seq(1:17),get_city)
    India_Fav_df<-as.data.frame(matrix(unlist(India_Fav), nrow=17, byrow=TRUE),stringsAsFactors = FALSE)
    
    ## Renames the columns and concvert the lat/lon to numerics
    colnames(India_Fav_df)<-c("City","lat","lng")
    
    India_Fav_df$lat<-as.numeric(India_Fav_df$lat)
    India_Fav_df$lng<-as.numeric(India_Fav_df$lng)
    
    ## Get the weather data for the desired cities
    get_weather_data<-function(i){
      weather_df<-get_current(
        India_Fav_df[i,1],
        units="metric")%>%flatten()
      return(weather_df[c("weather.icon","coord.lat","coord.lon","weather.main",
                          "name","main.pressure","main.humidity","main.temp","wind.speed")])%>%data.frame()
      
    }
  
    
    ## Get the colnames for the dataframe
    
   ## Names<-read.csv("names.csv",stringsAsFactors = FALSE)
    
    ## Convert the factor variable into character class
    
    
    India_Weather<-lapply(seq(1:17),get_weather_data)
    for(i in 1:17){
      India_Weather[i][[1]]$weather.icon=as.character(India_Weather[i][[1]]$weather.icon)}
    
    for(i in 1:17){
      India_Weather[i][[1]]$weather.main=as.character(India_Weather[i][[1]]$weather.main)}
    
    for(i in 1:17){
      India_Weather[i][[1]]$name=as.character(India_Weather[i][[1]]$name)}
    
    
    India_Weather_df<-as.data.frame(matrix(unlist(India_Weather), nrow=17, byrow=TRUE),stringsAsFactors = FALSE)
    colnames(India_Weather_df)<-Names$x
    
    ## COnvert the latitude and longitude into numerics
    
    India_Weather_df$coord_lat<-as.numeric(India_Weather_df$coord_lat)
    India_Weather_df$coord_lon<-as.numeric(India_Weather_df$coord_lon)
    India_Weather_df$main_temp<-as.numeric(India_Weather_df$main_temp)
    India_Weather_df$main_humidity<-as.numeric(India_Weather_df$main_humidity)
    
    ## Add Weather to the above map
    
    ## Icon
    
    popup_tpl <- paste0(
      "<b>{{name}}</b></br>",
      "{{coord_lon}}<, {{coord_lat}}</br>",
      "{{main_temp}}°C, ",
      "{{main_humidity}}, ",
      "<b>{{weather_main}}</b>")
    
    Map<-leaflet()%>%addProviderTiles(providers$Stamen.TerrainBackground)
    
    India_WeatherMap<-Map%>%add_weather(India_Weather_df,icon = India_Weather_df$weather_icon,template = popup_tpl)
    India_WeatherMap
    
  })
  
  output$plot1<-renderImage({
    
    ## Load the data for each  location and clean
    outfile <- tempfile(fileext='.gif')
    
    get_data<-function(i){
      location_df<-read.csv(i,stringsAsFactors = FALSE)
      clean_df<-location_df[,c(-1,-2,-5,-7,-9,-10,-11,-13,-15,-16,-17,-22,-23)]
      
      
      
      ## User defined fucntion to convert farenheit to celcius
      
      celcius<-function(x){
        c=(x-32)/1.8
        return(c)
      }
      
      
      
      ## Replace the "*" by gsub
      
      clean_df$MAX<-gsub("[*].*$","",clean_df$MAX)
      clean_df$MIN<-gsub("[*].*$","",clean_df$MIN)
      
      ## Convert the temperature to Celcius
      
      clean_df$TEMP<-celcius(clean_df$TEMP)
      clean_df$DEWP<-celcius(clean_df$DEWP)
      clean_df$MAX<-celcius(as.numeric(clean_df$MAX))
      clean_df$MIN<-celcius(as.numeric(clean_df$MIN))
      
      
      ## Convert the dates into character in order to split the coloumn into dd mm yy columns
      clean_df$YEARMODA<-as.Date(as.character(clean_df$YEARMODA),format="%Y%m%d")
      
      list<-strsplit(as.character(clean_df$YEARMODA),"-")
      ## Convert the list intok dataframe
      library(plyr)
      Date<-ldply(list)
      colnames(Date)<-c("Year","Month","Day")
      
      ## Column bind with the main dataframe
      clean_df<-cbind(clean_df,Date)
      ## Change the Date to numeric
      clean_df$Year=as.numeric(clean_df$Year)
      
      
      ## Return the final Dataset
      return(clean_df)
    }
    
    ## Get the clean data
  
    Station<-get_data(as.character(paste(input$station,".csv",sep="")))
    
      
    
    ## gganimate
    p2<-ggplot(Station, aes(x=Month, y=TEMP, color=factor(Month))) +
      geom_boxplot()+
      geom_boxplot(aes(frame = factor(Year)))+
      xlab("Months")+
      ylab("Average Temperature")+
      ggtitle("Average Temperature for last 4 Years")+
      theme(panel.border = element_blank(),
            panel.background = element_rect(fill=as.character(paste(input$color))),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            plot.title = element_text(hjust = 0.5,size=18,colour="indianred4"),
            axis.line = element_line(colour = "black"))+
      theme(legend.position="none")
    
    
    
    gganimate(p2,"outfile.gif")
    
    
    # Return a list containing the filename
    list(src = "outfile.gif",
         contentType = 'image/gif',
          width = 500,
          height = 450
         # alt = "This is alternate text"
    )}, 
    
    
    deleteFile = TRUE)
  
  
  output$yarrplot1<-renderPlot({
    
    rainfall<-read.csv("rainfall.csv",stringsAsFactors = FALSE)
    colnames(rainfall)<-c("Division","YEAR","Winter","Summer","Monsoon","Fall")
    ## Divide the dataset into pre and post 1950
    
    data<-if(input$Rainfall=="pre 1950"){
      rainfall[rainfall$YEAR<=1950,]
    }
    else{rainfall[rainfall$YEAR>1950,]}
    
    
    district<-group_by(data,Division)
    
    
    
    data_modified<-summarize(district,count=n(),Winter_Rainfall=mean(Winter,na.rm=TRUE),
                                 Summer_Rainfall=mean(Summer,na.rm=TRUE),
                                 Monsoon_Rainfall=mean(Monsoon,na.rm=TRUE),
                                 Fall_Rainfall=mean(Fall,na.rm=TRUE))
    
    ## Add new column of Yearly average
    data_modified<-mutate(data_modified,Yearly_Avg=(Winter_Rainfall+Summer_Rainfall+Monsoon_Rainfall+Fall_Rainfall)/4)
    data_sorted<-data_modified[order(-data_modified$Yearly_Avg),]
    data_sorted<-data_sorted[1:5,]
    
    list<-as.list(data_sorted)
    target<-c(list$Division[1],list$Division[2],
              list$Division[3],list$Division[4],
              list$Division[5])
    final_data=filter(rainfall,Division %in% target)
    
    ## pirateplot
    
    
    pirateplot(formula = as.formula(paste(input$Season, "~ Division")), 
               xlab="Indian Sub Terrain (Top 4)",
               ylab = "Rainfall in mm",
               data = final_data,
               main = "Pirateplot for top 4  Indian sub terrain pre 1950s",
               pal="pony",
               theme=1
               )
    
    
  })
  
  
  output$plot2<-renderImage({
    
    source("script.R")
    ## Create temporary gif file
    outfile1 <- tempfile(fileext='.gif')
    
## Load the saved India Map
    
    ## saveRDS(LocationMap, file = "Map")
     #x <- readRDS("Map")
     ##ggmap(x)
  
    ## Get the data loaded
    
    
    kol_data<-get_data("Kolkata.csv")
    Leh_data<-get_data("Leh.csv")
    Bikaner_data<-get_data("Bikaner.csv")
    Dehradun_data<-get_data("Dehradun.csv")
    Gangtok_data<-get_data("Gangtok.csv")
    Bangalore_data<-get_data("Bangalore.csv")
    
    # Get the latitude and longitude for the location and merge it with the dataset
    
    kol<-mutate(kol_data,City="Kolkata",lat=India_Fav_df[9,2],lon=India_Fav_df[9,3])
    Leh<-mutate(Leh_data,City="Leh",lat=India_Fav_df[1,2],lon=India_Fav_df[1,3])
    Bikaner<-mutate(Bikaner_data,City="Bikaner",lat=India_Fav_df[2,2],lon=India_Fav_df[2,3])
    Dehradun<-mutate(Dehradun_data,City="Dehradun",lat=30.31649,lon=78.03129)
    Gangtok<-mutate(Gangtok_data,City="Gangtok",lat=India_Fav_df[7,2],lon=India_Fav_df[7,3])
    Bangalore<-mutate(Bangalore_data,City="Bangalore",lat=India_Fav_df[11,2],lon=India_Fav_df[11,3])
    
    ## Row bind all the cities together
    
    Master_data<-rbind(kol,Leh,Bikaner,Dehradun,Gangtok,Bangalore)
    
    ## Get mean temperature
    
    Mean_Temp_India<-mean(Master_data$TEMP)
    
    ## Get the regions with temperature higher than average temperature
    
    Region_AboveMean<-Master_data[Master_data$TEMP>=Mean_Temp_India,]
    
    ## Get the Region with temperature lower than average temperature
    
    Region_BelowMean<-Master_data[Master_data$TEMP<Mean_Temp_India,]
    
    ## Plot it on the map with the help of ggplot
    
    
    p3 <- ggmap(x) + geom_point(data = if(input$Temperature=="Above Mean Temp of 4 years"){Region_AboveMean}else{Region_BelowMean}, 
                                  aes(x = lon, y = lat,size=30,color=City, 
                                      frame = Year))+
      
      ggtitle("Average Temperature for last 4 Years")+
      theme(panel.border = element_blank(),
            panel.background = element_rect(fill=as.character(paste(input$color))),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            plot.title = element_text(hjust = 0.5,size=18,colour="indianred4"),
            axis.line = element_line(colour = "black"))+
      theme(legend.position="none")
    
  
    gganimate(p3,"outfile1.gif")
    
    
    # Return a list containing the filename
    list(src = "outfile1.gif",
         contentType = 'image/gif',
         width = 500,
         height = 400
         # alt = "This is alternate text"
    )}, 
    
    
    deleteFile = TRUE)

  
})



```


##Final Thoughts

I hope you enjoyed this application. If you are interested then send me an email to richie.chatterjee31@gmail.com to work together for more complex application.



